---
title: "Demonstrating package & data setup and handling of sf objects: a case with read_soilmap()"
author: "Floris Vanderhaeghe"
date: "2020-03-30"
bibliography: references.bib
output: 
  rmarkdown::html_vignette:
    pandoc_args:
        - --csl
        - research-institute-for-nature-and-forest.csl
vignette: >
  %\VignetteIndexEntry{022. Demonstrating package & data setup and handling of sf objects: a case with read_soilmap()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

_Note: for more information on data storage and locations, see `vignette("v020_datastorage")`._
_The below workflow is an application and demonstrates current setup from R (output is mostly not included)._

## Minimal setup for using `read_soilmap()`

```{r}
library(remotes)
```

### Install and load `n2khab`

```{r}
install_github("inbo/n2khab",
               build_vignettes = TRUE,
               upgrade = TRUE)
```

```{r eval=TRUE}
library(n2khab)
```

### Create the `n2khab_data` directory with 2 subdirectories (in the working directory)

```{r}
n2khab_data_path <- fileman_folders()
```

You can also provide a custom root 'path' to `fileman_folders()`, but then take a location that is a (grand-grand-...) parent of the current working directory.
This is because reading functions of `n2khab` will by default use the first `n2khab_data` directory they can find, starting from the working directory and sequentially climbing up one directory level at a time.

### Download the `soilmap_simple` data source in the correct place

The `soilmap_simple` data source is a GeoPackage, provided and documented at
[Zenodo](https://doi.org/10.5281/zenodo.3732903).
It belongs to the collection of processed data ([Zenodo-link](https://zenodo.org/communities/n2khab-data-processed)).

```{r}
soilmap_simple_path <- file.path(n2khab_data_path, "20_processed/soilmap_simple")
dir.create(soilmap_simple_path)
download_zenodo(doi = "10.5281/zenodo.3732903",
                path = soilmap_simple_path)
```

At some time in future, the download will be performed automatically by `read_soilmap()` (if the `soilmap_simple` data source is missing).

## Explore the `soilmap_simple` data source!

```{r}
library(knitr)
library(tidyverse)
library(sf)
library(mapview)
library(plotly)
library(units)
```

Inspect documentation of `read_soilmap()`:

```{r}
?read_soilmap
```

Read the data source and inspect its contents:

```{r paged.print = FALSE}
sm_simple <- read_soilmap()
sm_simple
#> Simple feature collection with 270550 features and 10 fields
#> geometry type:  MULTIPOLYGON
#> dimension:      XY
#> bbox:           xmin: 22265.45 ymin: 153062.6 xmax: 258872.2 ymax: 244027.9
#> epsg (SRID):    31370
#> proj4string:    +proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013 +y_0=5400088.438 +ellps=intl +towgs84=-106.8686,52.2978,-103.7239,0.3366,-0.457,1.8422,-1.2747 +units=m +no_defs
#> # A tibble: 270,550 x 11
#>    bsm_poly_id bsm_region bsm_ge_coastalp… bsm_mo_soilunit… bsm_mo_substr
#>  *       <dbl> <fct>      <lgl>            <fct>            <fct>        
#>  1      165740 Kunstmati… TRUE             OB               <NA>         
#>  2      176046 Kunstmati… TRUE             OB               <NA>         
#>  3      185239 Zandleems… FALSE            Ldc              <NA>         
#>  4      162400 Kunstmati… TRUE             OB               <NA>         
#>  5      173971 Kunstmati… TRUE             OB               <NA>         
#>  6      173087 Zandleems… FALSE            Ldp              <NA>         
#>  7      199453 Zandleems… FALSE            Lep              <NA>         
#>  8      176922 Zandleems… FALSE            Ldc              <NA>         
#>  9      227861 Zandleems… FALSE            Abp(c)           <NA>         
#> 10      185390 Zandleems… FALSE            Lca              <NA>         
#> # … with 270,540 more rows, and 6 more variables: bsm_mo_tex <fct>,
#> #   bsm_mo_drain <fct>, bsm_mo_prof <fct>, bsm_mo_parentmat <fct>,
#> #   bsm_mo_profvar <fct>, geom <MULTIPOLYGON [m]>
```

```{r}
glimpse(sm_simple)
#> Observations: 270,550
#> Variables: 11
#> $ bsm_poly_id         <dbl> 165740, 176046, 185239, 162400, 173971, 173087, 1…
#> $ bsm_region          <fct> Kunstmatige gronden, Kunstmatige gronden, Zandlee…
#> $ bsm_ge_coastalplain <lgl> TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, FALSE, FALS…
#> $ bsm_mo_soilunitype  <fct> OB, OB, Ldc, OB, OB, Ldp, Lep, Ldc, Abp(c), Lca, …
#> $ bsm_mo_substr       <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, N…
#> $ bsm_mo_tex          <fct> NA, NA, L, NA, NA, L, L, L, A, L, NA, L, L, NA, N…
#> $ bsm_mo_drain        <fct> NA, NA, d, NA, NA, d, e, d, b, c, NA, d, c, NA, N…
#> $ bsm_mo_prof         <fct> NA, NA, c, NA, NA, p, p, c, p, a, NA, c, a, NA, N…
#> $ bsm_mo_parentmat    <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, N…
#> $ bsm_mo_profvar      <fct> NA, NA, NA, NA, NA, NA, NA, NA, (c), NA, NA, NA, …
#> $ geom                <MULTIPOLYGON [m]> MULTIPOLYGON (((204667.9 19..., MULT…
```

How many polygons are available per region?

```{r}
sm_simple %>%
    st_drop_geometry %>%
    count(bsm_region)
```

Wat is the average polygon area per region?

```{r}
sm_simple %>%
    mutate(area = st_area(.)) %>%
    st_drop_geometry %>%
    group_by(bsm_region) %>%
    summarise(mean_area = mean(area))
```

Plot polygons of the 'Zwin' region:

```{r}
zwin_map <-
    sm_simple %>%
    filter(bsm_region == "Zwin") %>%
    ggplot(aes(fill = bsm_mo_tex)) +
    geom_sf()
```

- with WGS84 graticule (though plotted in Belgian Lambert 72 = CRS 31370):

```{r}
zwin_map
```

- with Belgian Lambert 72 graticule:

```{r}
zwin_map + coord_sf(datum = st_crs(31370))
```

- make an interactive map of it:

```{r}
zwin_map %>% ggplotly()
```

- show data on top of [OpenTopoMap](https://opentopomap.org/about)'s Open Streetmap visualization:

```{r}
sm_simple %>%
        filter(bsm_region == "Zwin") %>%
        mutate(bsm_mo_tex = as.character(bsm_mo_tex)) %>%
        mapview(zcol = "bsm_mo_tex",
                alpha.region = 0.2,
                map.types = "OpenTopoMap")
```

Clicking a feature on the above generated map reveals all attributes.

## Access more information: the `soilmap` data source

### Download the `soilmap` data source in the correct place

The digital soil map of the Flemish Region is published
[at DOV](https://www.dov.vlaanderen.be/geonetwork/srv/dut/catalog.search#/metadata/5c129f2d-4498-4bc3-8860-01cb2d513f8f)
(Databank Ondergrond Vlaanderen).
The `soilmap` data source is the (renamed) shapefile, stored (with versioning) 
[at Zenodo](https://doi.org/10.5281/zenodo.3387007) in order
to support the `read_soilmap()` function and to sustain long-term workflow
reproducibility.
The `soilmap` data source belongs to the raw data collection ([Zenodo-link](https://zenodo.org/communities/n2khab-data-raw)).

```{r}
soilmap_path <- file.path(n2khab_data_path, "10_raw/soilmap")
dir.create(soilmap_path)
download_zenodo(doi = "10.5281/zenodo.3387008",
                path = soilmap_path,
                parallel = TRUE)
```

At some time in future, the download will be performed automatically by `read_soilmap()` (if the `soilmap` data source is missing).

### Explore the soilmap data source

Read the data source -- this takes a while (large dataset) -- and inspect its contents:

```{r}
sm <- read_soilmap(use_processed = FALSE)
sm
#> Simple feature collection with 270550 features and 37 fields
#> geometry type:  MULTIPOLYGON
#> dimension:      XY
#> bbox:           xmin: 22265.45 ymin: 153062.6 xmax: 258872.2 ymax: 244027.9
#> epsg (SRID):    31370
#> proj4string:    +proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013 +y_0=5400088.438 +ellps=intl +towgs84=-106.8686,52.2978,-103.7239,0.3366,-0.457,1.8422,-1.2747 +units=m +no_defs
#> # A tibble: 270,550 x 38
#>    bsm_poly_id bsm_map_id bsm_region bsm_ge_coastalp… bsm_ge_region bsm_legend
#>  *       <dbl> <fct>      <fct>      <lgl>            <fct>         <fct>     
#>  1      165740 61E        Kunstmati… TRUE             <NA>          Antropoge…
#>  2      176046 78W        Kunstmati… TRUE             <NA>          Antropoge…
#>  3      185239 95W        Zandleems… FALSE            <NA>          Vochtig z…
#>  4      162400 75E        Kunstmati… TRUE             <NA>          Antropoge…
#>  5      173971 63W        Kunstmati… TRUE             <NA>          Antropoge…
#>  6      173087 64W        Zandleems… FALSE            <NA>          Vochtig z…
#>  7      199453 98E        Zandleems… FALSE            <NA>          Nat zandl…
#>  8      176922 81W        Zandleems… FALSE            <NA>          Vochtig z…
#>  9      227861 88W        Zandleems… FALSE            <NA>          Droge leem
#> 10      185390 95W        Zandleems… FALSE            <NA>          Vochtig z…
#> # … with 270,540 more rows, and 32 more variables: bsm_legend_title <fct>,
#> #   bsm_legend_explan <fct>, bsm_soiltype_id <dbl>, bsm_soiltype <fct>,
#> #   bsm_soiltype_region <fct>, bsm_soilseries <fct>,
#> #   bsm_soilseries_explan <fct>, bsm_mo_soilunitype <fct>, bsm_mo_substr <fct>,
#> #   bsm_mo_substr_explan <fct>, bsm_mo_tex <fct>, bsm_mo_tex_explan <fct>,
#> #   bsm_mo_drain <fct>, bsm_mo_drain_explan <fct>, bsm_mo_prof <fct>,
#> #   bsm_mo_prof_explan <fct>, bsm_mo_parentmat <fct>,
#> #   bsm_mo_parentmat_explan <fct>, bsm_mo_profvar <fct>,
#> #   bsm_mo_profvar_explan <fct>, bsm_mo_phase <fct>, bsm_ge_substr <fct>,
#> #   bsm_ge_substr_explan <fct>, bsm_ge_series <fct>,
#> #   bsm_ge_series_explan <fct>, bsm_ge_subseries <fct>,
#> #   bsm_ge_subseries_explan <fct>, bsm_map_url <fct>, bsm_book_url <fct>,
#> #   bsm_detailmap_url <fct>, bsm_profloc_url <fct>, geometry <MULTIPOLYGON [m]>
```

```{r}
glimpse(sm)
```

Extract features that belong to the 'Middellandpolders' region:

```{r}
sm_mp <-
    sm %>%
    filter(bsm_region == "Middellandpolders")
dim(sm_mp)
#> [1] 3991   38
```

- make an interactive map (on top of Wikimedia's Open Streetmap visualization):

```{r}
sm_mp %>%
    mutate(bsm_ge_series = as.character(bsm_ge_series)) %>%
    mapview(zcol = "bsm_ge_series",
            alpha.region = 0.2,
            map.types = "Wikimedia",
            alpha = 0)
```

- calculate surface area (ha) and polygon count per `bsm_ge_series`:

```{r results='asis', paged.print=FALSE}
sm_mp %>%
    mutate(area = st_area(.) %>% set_units("ha")) %>%
    st_drop_geometry %>%
    group_by(bsm_ge_series, bsm_ge_series_explan) %>%
    summarise(area = sum(area) %>% round(2),
              nr_polygons = n()) %>%
    arrange(desc(area)) %>%
    kable
```

|bsm_ge_series |bsm_ge_series_explan                                                                                                                                           |         area| nr_polygons|
|:-------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------|------------:|-----------:|
|D             |overdekte kreekruggronden - Middelland materiaal over Oudland kreekrugmateriaal                                                                                | 9126.95 [ha]|        1387|
|E             |dekkleigronden - meer dan 100 cm Duinkerken III-klei                                                                                                           | 7160.28 [ha]|         761|
|F             |overdekte poelgronden en overdekte oude kleiplaatgronden met storende laag op geringe diepte                                                                   | 4848.24 [ha]|        1110|
|P             |overdekte Pleistocene gronden - gebroken poldermateriaal op Pleistoceen zand                                                                                   | 3506.81 [ha]|         582|
|G             |geulgronden - meer dan 100 cm zware Duinkerken III-klei in laaggelegen geulen                                                                                  |  815.64 [ha]|          72|
|A             |kreekruggronden - slibhoudend zand tot klei overgaand naar lichter materiaal                                                                                   |  308.98 [ha]|          27|
|A             |kreekruggronden - slibhoudend zand tot klei overgaand naar lichter materiaal + overdekte kreekrugronden - Middellland materiaal over Oudland kreekrugmateriaal |  230.75 [ha]|          33|
|M             |gronden van de Lage Moeren                                                                                                                                     |  163.11 [ha]|          19|

- calculate surface area (ha) and polygon count per soilseries where `bsm_ge_series == "D"`:

```{r results='asis', paged.print=FALSE}
sm_mp %>%
    filter(bsm_ge_series == "D") %>%
    mutate(area = st_area(.) %>% set_units("ha")) %>%
    st_drop_geometry %>%
    group_by(bsm_soilseries, bsm_soilseries_explan) %>%
    summarise(area = sum(area) %>% round(2),
              nr_polygons = n()) %>%
    arrange(desc(area)) %>%
    kable
```

|bsm_soilseries |bsm_soilseries_explan                                            |         area| nr_polygons|
|:--------------|:----------------------------------------------------------------|------------:|-----------:|
|m.D5           |Overdekte kreekruggronden (Middellandpolders)                    | 4628.42 [ha]|         642|
|m.Dk5          |Overdekte kreekruggronden - klei (Middellandpolders)             | 1287.51 [ha]|         123|
|m.Dl5          |Overdekte kreekruggronden - slibhoudend zand (Middellandpolders) | 1070.21 [ha]|         105|
|m.D4           |Overdekte kreekruggronden (Middellandpolders)                    |  843.91 [ha]|         259|
|m.D2           |Overdekte kreekruggronden (Middellandpolders)                    |  392.31 [ha]|          75|
|m.Dl6          |Overdekte kreekruggronden - slibhoudend zand (Middellandpolders) |  170.04 [ha]|          37|
|m.D5l          |Overdekte kreekruggronden (Middellandpolders)                    |  169.59 [ha]|          16|
|m.Dl4          |Overdekte kreekruggronden - slibhoudend zand (Middellandpolders) |  159.95 [ha]|          29|
|m.Dl2          |Overdekte kreekruggronden - slibhoudend zand (Middellandpolders) |  127.59 [ha]|          28|
|m.Dk6          |Overdekte kreekruggronden - klei (Middellandpolders)             |  115.02 [ha]|          19|
|m.D4l          |Overdekte kreekruggronden (Middellandpolders)                    |   85.74 [ha]|           6|
|m.D3           |Overdekte kreekruggronden (Middellandpolders)                    |   20.93 [ha]|           4|
|m.Df1          |Overdekte kreekruggronden - zware klei (Middellandpolders)       |   20.43 [ha]|          23|
|m.Dk4          |Overdekte kreekruggronden - klei (Middellandpolders)             |   15.08 [ha]|           6|
|m.D1           |Overdekte kreekruggronden (Middellandpolders)                    |   13.74 [ha]|           9|
|m.D5d          |Overdekte kreekruggronden (Middellandpolders)                    |    6.46 [ha]|           6|

- plot all features where `bsm_soilseries == "m.D5"`:

```{r}
sm_mp %>%
    filter(bsm_soilseries == "m.D5") %>%
    mapview(color = "red",
            alpha = 1,
            alpha.region = 0)
```

