---
title: "Using the GRTS data sources"
author: "Floris Vanderhaeghe"
date: "2019-05-29"
bibliography: references.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{030. Using the GRTS data sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(n2khab)
library(magrittr)
library(knitr)
datapath <- "../../data"
```

## Intro

With this vignette, you get acquainted with three functions that usually return a RasterLayer or a RasterBrick:

- `read_GRTSmh()`
- `read_GRTSmh_base4frac()`
- `read_GRTSmh_diffres()`

The data source `GRTSmaster_habitats`, provided and documented in [Zenodo](https://doi.org/10.5281/zenodo.2611233), is a monolayered GeoTIFF file covering the whole of Flanders and the Brussels Capital Region at a resolution of 32 m.
Its values are unique decimal integer ranking numbers from the GRTS algorithm applied to the Flemish and Brussels area.
Beware that not all GRTS ranking numbers are present in the data source, as the original GRTS raster has been clipped with the Flemish outer borders (i.e., not excluding the Brussels Capital Region).

The GRTS algorithm uses a quadrant-recursive, hierarchically randomized function that maps the unit square to the unit interval, resulting in a base-4 GRTS address for each location. The ranking numbers in `GRTSmaster_habitats` are base-10 numbers and follow the reverse hierarchical order: **each consecutive subset** of ranking numbers corresponds to a **spatially balanced** sample of locations. Hence, it allows **dynamical** sample sizes. More information on the GRTS algorithm can be found in @stevens_variance_2003 [-@stevens_spatially_2004] and in the [GRTS](https://github.com/ThierryO/grts) and [spsurvey](https://cran.r-project.org/web/packages/spsurvey) packages.


## Data sources

The following data sources are available:

- the raw data source `GRTSmaster_habitats`, discussed above and available at [Zenodo](https://doi.org/10.5281/zenodo.2611233)
- processed data sources:
    - `GRTSmh_brick`: 10-layered GeoTIFF with the decimal integer ranking numbers of 10 hierarchical levels (0 - 9) of the GRTS cell addresses, including the one from `GRTSmaster_habitats` (i.e. level 0; for more details see the [read_GRTSmh() documentation](../html/read_GRTSmh.html))
    - `GRTSmh_diffres`: file collection composed of nine monolayered GeoTIFF files plus a GeoPackage with six polygon layers. They provide the hierarchical levels 1 to 9 of the `GRTSmh_brick` data source at the corresponding spatial resolution, i.e. at lower resolutions than `GRTSmaster_habitats` (for more details see the [read_GRTSmh_diffres() documentation](../html/read_GRTSmh_diffres.html))
    - `GRTSmh_base4frac`: is like a mirror to `GRTSmaster_habitats`, holding the ranking numbers as base 4 fractions (for more details see the [read_GRTSmh_base4frac() documentation](../html/read_GRTSmh_base4frac.html))

For more information on data storage and locations, there is a [specific vignette](../doc/vign-020_datastorage.html).

## Get the data in R

### read_GRTSmh()

In the below R code, the recommendation has been followed to set the location of the `data` folder by a variable **`datapath`**.
See the [vignette on data storage](../doc/vign-020_datastorage.html) for more information.

`read_GRTSmh()` by default returns the `GRTSmaster_habitats` dataset:

```{r eval=FALSE}
read_GRTSmh(datapath)
#> class       : RasterLayer 
#> dimensions  : 2843, 7401, 21041043  (nrow, ncol, ncell)
#> resolution  : 32, 32  (x, y)
#> extent      : 22029.59, 258861.6, 153054.1, 244030.1  (xmin, xmax, ymin, ymax)
#> coord. ref. : +proj=lcc +lat_1=49.8333339 +lat_2=51.16666723333333 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.01256 +y_0=5400088.4378 +ellps=intl +units=m +no_defs 
#> data source : [...]/n2khab-preprocessing/data/10_raw/GRTSmaster_habitats/GRTSmaster_habitats.tif 
#> names       : GRTSmaster_habitats 
#> values      : 1, 67108857  (min, max)
```

With the argument `brick = TRUE` however, you will get the `GRTSmh_brick` data source, i.e. `GRTSmaster_habitats` plus 9 extra layers:

```{r eval=FALSE}
read_GRTSmh(datapath, brick = TRUE)
#> class       : RasterBrick 
#> dimensions  : 2843, 7401, 21041043, 10  (nrow, ncol, ncell, nlayers)
#> resolution  : 32, 32  (x, y)
#> extent      : 22029.59, 258861.6, 153054.1, 244030.1  (xmin, xmax, ymin, ymax)
#> coord. ref. : +proj=lcc +lat_1=49.8333339 +lat_2=51.16666723333333 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.01256 +y_0=5400088.4378 +ellps=intl +units=m +no_defs 
#> data source : [...]/n2khab-preprocessing/data/20_processed/GRTSmh_brick/GRTSmh_brick.tif 
#> names       :   level0,   level1,   level2,   level3,   level4,   level5,   level6,   level7,   level8,   level9 
#> min values  :        1,        1,        1,        1,        1,        1,        1,        1,        1,        1 
#> max values  : 67108857, 16777209,  4194297,  1048569,   262137,    65529,    16377,     4089,     1017,      253
```

The higher-level ranking numbers of the RasterBrick allow spatially balanced samples at lower spatial resolution than that of 32 m, and can also be used for aggregation purposes.
The provided hierarchical levels correspond to the resolution vector `32 * 2^(0:9)` (minimum: 32 meters, maximum: 16384 meters), with the corresponding RasterBrick layers named as `level0` to `level9`.

### read_GRTSmh_diffres()

`read_GRTSmh_diffres()` by default returns one RasterLayer from the `GRTSmh_diffres` data source, i.e. with the GRTS ranking numbers of the user-specified hierarchical level.
This is done _at the corresponding spatial resolution_ of the GRTS algorithm, which is the fundamental distinction from `read_GRTSmh(brick = TRUE)`.

The resolutions of each level are the following (in meters):

```{r warning = FALSE, echo = FALSE}
data.frame(level = 1:9,
           resolution = 32 * 2 ^ (1:9)) %>% 
  kable(align = "r")
```

An example with level 5:

```{r eval=FALSE}
read_GRTSmh_diffres(datapath, level = 5)
#> class      : RasterLayer 
#> dimensions : 89, 232, 20648  (nrow, ncol, ncell)
#> resolution : 1024, 1024  (x, y)
#> extent     : 22030, 259598, 153054, 244190  (xmin, xmax, ymin, ymax)
#> crs        : +proj=lcc +lat_1=49.8333339 +lat_2=51.16666723333333 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.01256 +y_0=5400088.4378 +ellps=intl +units=m +no_defs 
#> source     : [...]/n2khab-preprocessing/data/20_processed/GRTSmh_diffres/GRTSmh_diffres.5.tif 
#> names      : level5 
#> values     : 1, 65529  (min, max)
```

Alternatively, a dissolved, polygonized variant of the corresponding `GRTSmh_brick` level can be returned as an `sf` object.
In order not to inflate the data source, this was only made available for levels 4 to 9.

```{r eval=FALSE}
read_GRTSmh_diffres(datapath, level = 5, polygon = TRUE)
#> Simple feature collection with 13791 features and 1 field
#> geometry type:  POLYGON
#> dimension:      XY
#> bbox:           xmin: 22029.59 ymin: 153054.1 xmax: 258861.6 ymax: 244030.1
#> epsg (SRID):    NA
#> proj4string:    +proj=lcc +lat_1=49.8333339 +lat_2=51.16666723333333 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.01256 +y_0=5400088.4378 +ellps=intl +units=m +no_defs
#> First 10 features:
#>    value                           geom
#> 1  23390 POLYGON ((178093.6 244030.1...
#> 2  56158 POLYGON ((178701.6 243646.1...
#> 3  23134 POLYGON ((177581.6 243870.1...
#> 4  60254 POLYGON ((179757.6 243454.1...
#> 5   6750 POLYGON ((176621.6 243390.1...
#> 6  60254 POLYGON ((180333.6 243230.1...
#> 7  58718 POLYGON ((176653.6 243166.1...
#> 8  52318 POLYGON ((177677.6 243166.1...
#> 9   3166 POLYGON ((178701.6 243166.1...
#> 10 15454 POLYGON ((179725.6 243166.1...
```


### read_GRTSmh_base4frac()

Its use is just to return the base-4-fraction-converted `GRTSmaster_habitats` as a RasterLayer:

```{r eval=FALSE}
options(scipen = 999, digits = 15)
read_GRTSmh_base4frac(datapath)
#> class       : RasterLayer 
#> dimensions  : 2843, 7401, 21041043  (nrow, ncol, ncell)
#> resolution  : 32, 32  (x, y)
#> extent      : 22029.591973471, 258861.591973471, 153054.113583292, 244030.113583292  (xmin, xmax, ymin, ymax)
#> coord. ref. : +proj=lcc +lat_1=49.8333339 +lat_2=51.16666723333333 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.01256 +y_0=5400088.4378 +ellps=intl +units=m +no_defs 
#> data source : [...]/n2khab-preprocessing/data/20_processed/GRTSmh_base4frac/GRTSmh_base4frac.tif 
#> names       : GRTSmh_base4frac 
#> values      : 0.0000000000001, 0.3333333333321  (min, max)
```

Note that the used options are necessary when treating these base-4-fraction GRTS addresses as characters; otherwise scientific notations will be used.

Also, be warned that R does not actually regard the values as base 4, but as base 10. [^base4]

[^base4]: So, what really matters is only the notation with many digits, to be _regarded_ as a base 4 fraction (and hence, handling it as a character in conversions is often necessary).
The `n2khab` package also exports a `dec_to_base4frac()` and a `base4frac_to_dec()` function in its namespace.
These functions will be relevant if you need to do such conversions yourself, and they are used in the code to generate the processed data sources.


## References


