# Retrieving the list of current type codes from the reference googlesheet and splitting it in two n2khab data sources

## Initialize `namelist` with typeclasses

Set path and filename of the namelist:

```{r}
namelist_path <- "../../inst/textdata/namelist"
```


Reading from the source googlesheet (maintained by the MNE and [HabNorm](https://github.com/inbo/habnorm) projects):

```{r message=FALSE}
types_source <- gs_key("1dK0S1Tt3RlVEh4WrNF5N_-hn0OXiTUVOvsL2JRhSl78")
# gs_browse(types_gs)
source_df <-
    types_source %>%
    gs_read(ws = 1, 
            check.names = T,
            verbose = FALSE)
```

Writing the namelist a first time:

```{r message=FALSE}
dummy <- 
    str_c(namelist_path, ".*") %>% 
    Sys.glob %>% 
    file.remove
source_df %>% 
    distinct(Groepering.niveau.1,
           Groepering.niveau.1.Eng) %>% 
    rename(nl = Groepering.niveau.1,
           en = Groepering.niveau.1.Eng) %>% 
    inner_join(tribble(
        ~code, ~en,
        "CH", "Coastal and halophytic habitats",
        "CD", "Coastal sand dunes", 
        "ID", "Inland dunes", 
        "FW", "Fresh and brackish water", 
        "HS", "Temperate heath and scrub", 
        "SS", "Sclerophyllous scrub", 
        "GR", "Natural and semi-natural grassland",
        "BMF", "Bogs, mires and fens", 
        "RC", "Rocky habitats and caves", 
        "FS", "Forest and scrub"
    )) %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    write_vc(namelist_path,
             sorting = c("lang", "code"))
```

## Append typenames to `namelist`

```{r}
source_df %>% 
    select(code = Code.geïntegreerd,
           nl_name = Volledige.naam,
           en_name = Volledige.naam.Eng,
           nl_shortname = Verkorte.naam,
           en_shortname = Verkorte.naam.Eng) %>% 
    gather(key = "key", 
           value = "value",
           contains("name")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    bind_rows(read_vc(namelist_path),
              .) %>% 
    write_vc(namelist_path)
```

## Append extra tags to `namelist`

```{r message=FALSE}
source_df %>% 
    distinct(nl = Groepering.niveau.2,
           en = Groepering.niveau.2.Eng) %>% 
    filter(!is.na(nl) |
               !is.na(en)
           ) %>% 
    inner_join(tribble(
        ~code, ~en,
        "DO", "Dike outward",
        "DIO", "Dike in-/outward", 
        "DI", "Dike inward", 
        "SW", "Standing water", 
        "RW", "Running water", 
        "CT", "Calcareous type"
    )) %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path),
              .) %>% 
    write_vc(namelist_path)
```

## Make `types`

Set path and filename of `types`:

```{r}
types_path <- "../../inst/textdata/types"
```

Write the data source `types`:

```{r}
en_codelist <- 
    read_vc(namelist_path) %>% 
    filter(lang == "en") %>% 
    select(code, name)
source_df %>%
    select(type = Code.geïntegreerd,
           main_type = Code,
           typeclass_en = Groepering.niveau.1.Eng,
           tag_1_en = Groepering.niveau.2.Eng) %>% 
    mutate(typelevel = ifelse(main_type == type, 
                              "main_type",
                              "subtype") %>% 
               factor) %>% 
    inner_join(en_codelist,
               by = c("typeclass_en" = "name")) %>% 
    rename(typeclass = code) %>% 
    left_join(en_codelist,
              by = c("tag_1_en" = "name")) %>% 
    select(type,
           typelevel,
           main_type,
           typeclass,
           tag_1 = code) %>% 
    mutate(typeclass = factor(typeclass, 
                              levels = (.$typeclass %>% unique)),
           tag_2 = NA %>% as.character,
           tag_3 = NA %>% as.character
           ) %>% 
    arrange(typeclass, 
            type) %>% 
    mutate(type = factor(type, levels = .$type),
           main_type = factor(main_type, 
                              levels = .$main_type %>% unique)
           ) %>% 
    write_vc(types_path,
             sorting = c("typeclass",
                         "type"),
             optimize = FALSE,
             strict = FALSE)
```



