# Retrieving the list of current type codes from the reference googlesheet and splitting it in two n2khab data sources

## Initialize `namelist` with typeclasses

Set path and filename of the namelist:

```{r}
namelist_path <- "inst/textdata/namelist"
root <- "../.."
```


Reading from the source googlesheet (maintained by Steven De Saeger):

```{r message=FALSE}
# gs4_browse(types_gs_id)
source_df <-
    types_gs_id %>%
    read_sheet(sheet = 1,
               .name_repair = make.names) %>% 
    mutate(across(starts_with("Code"), unlist))
```

Writing the namelist a first time:

```{r}
dummy <- 
    str_c(root, "/", namelist_path, ".*") %>% 
    Sys.glob %>% 
    file.remove
source_df %>% 
    distinct(Groepering.niveau.1,
           Groepering.niveau.1.Eng) %>% 
    rename(nl = Groepering.niveau.1,
           en = Groepering.niveau.1.Eng) %>% 
    inner_join(tribble(
        ~code, ~en,
        "CH", "Coastal and halophytic habitats",
        "CD", "Coastal sand dunes", 
        "ID", "Inland dunes", 
        "FW", "Fresh and brackish water", 
        "HS", "Temperate heath and scrub", 
        "SS", "Sclerophyllous scrub", 
        "GR", "Natural and semi-natural grassland",
        "BMF", "Bogs, mires and fens", 
        "RC", "Rocky habitats and caves", 
        "FS", "Forest and scrub"
    ), by = "en") %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    write_vc(namelist_path,
             root,
             sorting = c("lang", "code"))
```

## Append typenames to `namelist`

```{r}
source_df %>% 
    select(code = Code.geïntegreerd,
           nl_name = Volledige.naam,
           en_name = Volledige.naam.Eng,
           nl_shortname = Verkorte.naam,
           en_shortname = Verkorte.naam.Eng) %>% 
    gather(key = "key", 
           value = "value",
           contains("name")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

## Append extra labels to `namelist`

Information for tag 1:

```{r}
source_df %>% 
    distinct(nl = Groepering.niveau.2,
           en = Groepering.niveau.2.Eng) %>% 
    filter(!is.na(nl) |
               !is.na(en)
           ) %>% 
    inner_join(tribble(
        ~code, ~en,
        "DO", "Dike outward",
        "DIO", "Dike in-/outward", 
        "DI", "Dike inward", 
        "SW", "Standing water", 
        "RW", "Running water", 
        "CT", "Calcareous type"
    ), by = "en") %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Hydrological classes:

```{r}
type_hydroclass <- 
    read_vc("rawraw_data/vegwhh") %>% 
    rename(type = Vegcode,
           hydroclass = Waterhuishouding) %>% 
    group_by(type) %>% 
    mutate(count_hydrocl = n(),
           hydroclass = as.character(hydroclass)) %>% 
    ungroup %>% 
    spread(key = hydroclass, value = hydroclass) %>% 
    rename(L1 = "Droog tot vochtig",
           L2 = "Tijdelijk tot permanent nat",
           L3 = "Oppervlaktewater") %>% 
    unite(nl, L1, L2, L3, sep = " + ", na.rm = TRUE) %>% 
    select(-count_hydrocl)
# type_hydroclass %>% 
#     distinct(nl)
hc_names <- 
    tribble(
        ~code, ~nl_name, ~en_name, ~nl_shortname, ~en_shortname,
        "HC1", "Droog tot vochtig", "Dry to moist",
                "Droog", "Dry",
        "HC12", "Droog tot vochtig + Tijdelijk tot permanent nat", 
                "Dry to moist + Temporarily to permanently wet",
                "Droog + Nat", "Dry + Wet",
        "HC2", "Tijdelijk tot permanent nat", "Temporarily to permanently wet",
                "Nat", "Wet",
        "HC23", "Tijdelijk tot permanent nat + Oppervlaktewater", 
                "Temporarily to permanently wet + Surface water",
                "Nat + Oppervlaktewater", "Wet + Surface water",
        "HC3", "Oppervlaktewater", "Surface water",
        "Oppervlaktewater", "Surface water"
    )
type_hydroclass <- 
    type_hydroclass %>% 
    inner_join(hc_names %>% select(code, 
                                   nl_name),
               by = c("nl" = "nl_name")) %>% 
    select(type, code)
    
hc_names %>% 
    select(1:3) %>% 
    pivot_longer(cols = -code,
                 names_to = "lang",
                 names_pattern = "(..)_.+",
                 values_to = "name") %>% 
    inner_join(
        hc_names %>% 
        select(1, 4:5) %>% 
        pivot_longer(cols = -code,
                     names_to = "lang",
                     names_pattern = "(..)_.+",
                     values_to = "shortname"),
        by = c("code", "lang")
    ) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Groundwater and flood dependency:

```{r message=FALSE}
# gs4_browse(wdep_gs_id)
type_wdep <-
    wdep_gs_id %>%
    read_sheet(sheet = "Waterafhankelijke HT en RBB",
            skip = 2,
            col_names = FALSE) %>%  
    select(type = ...1, 
           groundw_dep = ...9,
           flood_dep = ...10) %>% 
    mutate(groundw_dep = case_when(str_detect(groundw_dep, "alle") ~ "GD2",
                                    str_detect(groundw_dep, "niet") ~ "GD0",
                                    TRUE ~ "GD1") %>% 
                        factor,
           flood_dep = case_when(type %in% subset(type_hydroclass, code == "HC3")$type ~ NA_character_,
                                 str_detect(flood_dep, "\\(.+\\)") ~ "FD2",
                                    str_detect(flood_dep, "niet") ~ "FD0",
                                    TRUE ~ "FD1") %>% 
                        factor)

wdep_names <- 
    tribble(
        ~code, ~nl_name, ~nl_shortname, ~en_name, ~en_shortname,
        "GD0", "niet grondwaterafhankelijk", "niet grondwaterafhankelijk", "groundwater independent", "groundwater independent",
        "GD1", "grondwaterafhankelijk op een deel van de locaties", "plaatselijk grondwaterafhankelijk", "groundwater dependent on a part of the locations", "locally groundwater dependent", 
        "GD2", "grondwaterafhankelijk op (bijna) alle locaties", "(bijna) overal grondwaterafhankelijk", "groundwater dependent on (almost) all locations", "(almost) everywhere groundwater dependent", 
        "FD0", "niet overstromingsafhankelijk", "niet overstromingsafhankelijk", "flood independent", "flood independent", 
        "FD1", "overstromingsafhankelijk op een deel van de locaties", "plaatselijk overstromingsafhankelijk", "flood dependent on a part of the locations", "locally flood dependent", 
        "FD2", "overstromingsafhankelijk op (bijna) alle locaties", "(bijna) overal overstromingsafhankelijk", "flood dependent on (almost) all locations", "(almost) everywhere flood dependent"
)

wdep_names %>% 
    select(1, 2, 4) %>% 
    pivot_longer(cols = -code,
                 names_to = "lang",
                 names_pattern = "(..)_.+",
                 values_to = "name") %>% 
    inner_join(
        wdep_names %>% 
        select(1, 3, 5) %>% 
        pivot_longer(cols = -code,
                     names_to = "lang",
                     names_pattern = "(..)_.+",
                     values_to = "shortname"),
        by = c("code", "lang")
    ) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Note that we have deliberately set flood dependency as `NA` for aquatic types, i.e. from hydrological class 'surface water' (code HC3).

## Make `types`

Set path and filename of `types`:

```{r}
types_path <- "inst/textdata/types"
```

Write the data source `types`:

```{r}
en_codelist <- 
    read_vc(namelist_path, root) %>% 
    filter(lang == "en") %>% 
    select(code, name)
source_df %>%
    select(type = Code.geïntegreerd,
           main_type = Code,
           typeclass_en = Groepering.niveau.1.Eng,
           tag_1_en = Groepering.niveau.2.Eng) %>% 
    mutate(typelevel = ifelse(main_type == type, 
                              "main_type",
                              "subtype") %>% 
               factor) %>% 
    inner_join(en_codelist,
               by = c("typeclass_en" = "name")) %>% 
    rename(typeclass = code) %>% 
    left_join(type_hydroclass %>% 
                  mutate(code = factor(code)) %>% 
                  rename(hydr_class = code),
              by = "type") %>% 
    left_join(type_wdep,
              by = "type") %>% 
    left_join(en_codelist,
              by = c("tag_1_en" = "name")) %>% 
    select(type,
           typelevel,
           main_type,
           typeclass,
           hydr_class,
           groundw_dep,
           flood_dep,
           tag_1 = code) %>% 
    mutate(typeclass = factor(typeclass, 
                              levels = (.$typeclass %>% unique)),
           tag_2 = NA %>% as.character,
           tag_3 = NA %>% as.character
           ) %>% 
    arrange(typeclass, 
            type) %>% 
    mutate(type = factor(type, levels = .$type),
           main_type = factor(main_type, 
                              levels = .$main_type %>% unique)
           ) %>% 
    write_vc(types_path,
             root,
             sorting = c("typeclass",
                         "type"),
             optimize = FALSE,
             strict = FALSE)
```



