# Retrieving the list of current type codes from the reference googlesheet and splitting it in two n2khab data sources

## Initialize `namelist` with typeclasses

Set path and filename of the namelist:

```{r}
namelist_path <- "inst/textdata/namelist"
root <- "../.."
```


Reading from the source googlesheet (maintained by Steven De Saeger):

```{r message=FALSE}
# the first time, run:
# gs_auth()
types_source <- gs_key("1dK0S1Tt3RlVEh4WrNF5N_-hn0OXiTUVOvsL2JRhSl78")
# gs_browse(types_source)
source_df <-
    types_source %>%
    gs_read(ws = 1, 
            check.names = T,
            verbose = FALSE)
```

Writing the namelist a first time:

```{r message=FALSE}
dummy <- 
    str_c(root, "/", namelist_path, ".*") %>% 
    Sys.glob %>% 
    file.remove
source_df %>% 
    distinct(Groepering.niveau.1,
           Groepering.niveau.1.Eng) %>% 
    rename(nl = Groepering.niveau.1,
           en = Groepering.niveau.1.Eng) %>% 
    inner_join(tribble(
        ~code, ~en,
        "CH", "Coastal and halophytic habitats",
        "CD", "Coastal sand dunes", 
        "ID", "Inland dunes", 
        "FW", "Fresh and brackish water", 
        "HS", "Temperate heath and scrub", 
        "SS", "Sclerophyllous scrub", 
        "GR", "Natural and semi-natural grassland",
        "BMF", "Bogs, mires and fens", 
        "RC", "Rocky habitats and caves", 
        "FS", "Forest and scrub"
    )) %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    write_vc(namelist_path,
             root,
             sorting = c("lang", "code"))
```

## Append typenames to `namelist`

```{r}
source_df %>% 
    select(code = Code.geïntegreerd,
           nl_name = Volledige.naam,
           en_name = Volledige.naam.Eng,
           nl_shortname = Verkorte.naam,
           en_shortname = Verkorte.naam.Eng) %>% 
    gather(key = "key", 
           value = "value",
           contains("name")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

## Append extra labels to `namelist`

Information for tag 1:

```{r message=FALSE}
source_df %>% 
    distinct(nl = Groepering.niveau.2,
           en = Groepering.niveau.2.Eng) %>% 
    filter(!is.na(nl) |
               !is.na(en)
           ) %>% 
    inner_join(tribble(
        ~code, ~en,
        "DO", "Dike outward",
        "DIO", "Dike in-/outward", 
        "DI", "Dike inward", 
        "SW", "Standing water", 
        "RW", "Running water", 
        "CT", "Calcareous type"
    )) %>% 
    gather(key = "lang", value = "name", en:nl) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Hydrological classes:

```{r}
type_hydroclass <- 
    read_vc("rawraw_data/vegwhh") %>% 
    rename(type = Vegcode,
           hydroclass = Waterhuishouding) %>% 
    group_by(type) %>% 
    mutate(count_hydrocl = n(),
           hydroclass = as.character(hydroclass)) %>% 
    ungroup %>% 
    spread(key = hydroclass, value = hydroclass) %>% 
    rename(L1 = "Droog tot vochtig",
           L2 = "Tijdelijk tot permanent nat",
           L3 = "Oppervlaktewater") %>% 
    unite(nl, L1, L2, L3, sep = " + ", na.rm = TRUE) %>% 
    select(-count_hydrocl)
# type_hydroclass %>% 
#     distinct(nl)
hc_names <- 
    tribble(
        ~code, ~nl_name, ~en_name, ~nl_shortname, ~en_shortname,
        "HC1", "Droog tot vochtig", "Dry to moist",
                "Droog", "Dry",
        "HC12", "Droog tot vochtig + Tijdelijk tot permanent nat", 
                "Dry to moist + Temporarily to permanently wet",
                "Droog + Nat", "Dry + Wet",
        "HC2", "Tijdelijk tot permanent nat", "Temporarily to permanently wet",
                "Nat", "Wet",
        "HC23", "Tijdelijk tot permanent nat + Oppervlaktewater", 
                "Temporarily to permanently wet + Surface water",
                "Nat + Oppervlaktewater", "Wet + Surface water",
        "HC3", "Oppervlaktewater", "Surface water",
        "Oppervlaktewater", "Surface water"
    )
type_hydroclass <- 
    type_hydroclass %>% 
    inner_join(hc_names %>% select(code, 
                                   nl_name),
               by = c("nl" = "nl_name")) %>% 
    select(type, code)
    
hc_names %>% 
    select(1:3) %>% 
    pivot_longer(cols = -code,
                 names_to = "lang",
                 names_pattern = "(..)_.+",
                 values_to = "name") %>% 
    inner_join(
        hc_names %>% 
        select(1, 4:5) %>% 
        pivot_longer(cols = -code,
                     names_to = "lang",
                     names_pattern = "(..)_.+",
                     values_to = "shortname")
    ) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```



## Make `types`

Set path and filename of `types`:

```{r}
types_path <- "inst/textdata/types"
```

Write the data source `types`:

```{r}
en_codelist <- 
    read_vc(namelist_path, root) %>% 
    filter(lang == "en") %>% 
    select(code, name)
source_df %>%
    select(type = Code.geïntegreerd,
           main_type = Code,
           typeclass_en = Groepering.niveau.1.Eng,
           tag_1_en = Groepering.niveau.2.Eng) %>% 
    mutate(typelevel = ifelse(main_type == type, 
                              "main_type",
                              "subtype") %>% 
               factor) %>% 
    inner_join(en_codelist,
               by = c("typeclass_en" = "name")) %>% 
    rename(typeclass = code) %>% 
    left_join(type_hydroclass %>% 
                  mutate(code = factor(code)) %>% 
                  rename(hydr_class = code),
              by = "type") %>% 
    left_join(en_codelist,
              by = c("tag_1_en" = "name")) %>% 
    select(type,
           typelevel,
           main_type,
           typeclass,
           hydr_class,
           tag_1 = code) %>% 
    mutate(typeclass = factor(typeclass, 
                              levels = (.$typeclass %>% unique)),
           tag_2 = NA %>% as.character,
           tag_3 = NA %>% as.character
           ) %>% 
    arrange(typeclass, 
            type) %>% 
    mutate(type = factor(type, levels = .$type),
           main_type = factor(main_type, 
                              levels = .$main_type %>% unique)
           ) %>% 
    write_vc(types_path,
             root,
             sorting = c("typeclass",
                         "type"),
             optimize = FALSE,
             strict = FALSE)
```



