# Retrieving the current list of environmental pressures from the reference googlesheet and moving data to two n2khab data sources

## Append extra names to `namelist`

Reading from the source googlesheet:

```{r message=FALSE}
ep_source <- gs_key("1PH6InqJk0ijQF_N7v7IZjarijlqHRBCKhTbDn44skZU")
source_df <-
    ep_source %>%
    gs_read(ws = gs_ws_ls(ep_source) %>%
                 tail(1), 
            check.names = T,
            verbose = FALSE)
```

Adding multilingual EP names and abbreviations to `namelist`:

```{r}
source_df %>% 
    select(code = Environmental_pressure_code,
           nl_name = Environmental_pressure_name_NL,
           en_name = Environmental_pressure_name_EN,
           nl_shortname = Environmental_pressure_abbrev_NL,
           en_shortname = Environmental_pressure_abbrev_EN) %>% 
    gather(key = "key", 
           value = "value",
           matches(".+name|.+explanation|.+epclass")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Adding explanations to `namelist`:

```{r}
source_df %>% 
    select(code = Environmental_pressure_code,
           nl_name = Explanation_NL,
           en_name = Explanation_EN) %>% 
    mutate(code = str_c(code, "_explan")) %>% 
    gather(key = "key", 
           value = "value",
           contains("name")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    filter(!is.na(name)) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Adding ep_classes to `namelist`:

```{r}
source_df %>% 
    select(code = Environmental_pressure_code,
           nl_name = Environmental_pressure_class_NL,
           en_name = Environmental_pressure_class_EN) %>% 
    gather(key = "key", 
           value = "value",
           contains("name")) %>% 
    separate(key, c("lang", "key")) %>% 
    spread(key = key,
           value = value) %>% 
    filter(!is.na(name)) %>% 
    mutate(code = str_c("ep_class_",
                        str_match(code, "ep_(..).*")[,2]
                        )
           ) %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```


## Generate the `env_pressures` data source

Set path and filename of `env_pressures`:

```{r}
ep_path <- "inst/textdata/env_pressures"
```

We generate it manually because of the difficulties with cells spanning several rows in the googlesheet:

```{r}
tribble(
    ~ep_code, ~ep_class, ~explanation,
    "ep_011", "ep_class_01", "ep_011_explan", 
    "ep_012", "ep_class_01", "ep_012_explan", 
    "ep_013", "ep_class_01", "ep_012_explan", 
    "ep_014", "ep_class_01", "ep_014_explan", 
    "ep_015", "ep_class_01", "ep_015_explan", 
    "ep_03.1", "ep_class_03", "ep_03.1_explan", 
    "ep_03.2", "ep_class_03", "ep_03.1_explan", 
    "ep_03.3", "ep_class_03", "ep_03.1_explan", 
    "ep_03.4", "ep_class_03", "ep_03.1_explan", 
    "ep_04.1", "ep_class_04", "ep_04.1_explan", 
    "ep_04.2", "ep_class_04", "ep_04.1_explan", 
    "ep_04.3", "ep_class_04", "ep_04.1_explan", 
    "ep_05.1", "ep_class_05", "ep_05.1_explan", 
    "ep_05.2", "ep_class_05", "ep_05.1_explan", 
    "ep_061", "ep_class_06", "ep_061_explan", 
    "ep_062", "ep_class_06", "ep_061_explan", 
    "ep_063", "ep_class_06", "ep_063_explan", 
    "ep_064", "ep_class_06", "ep_063_explan", 
    "ep_065", "ep_class_06", "ep_065_explan", 
    "ep_066", "ep_class_06", "ep_065_explan", 
    "ep_07.1", "ep_class_07", "ep_07.1_explan", 
    "ep_07.2", "ep_class_07", "ep_07.1_explan", 
    "ep_07.3", "ep_class_07", "ep_07.1_explan", 
    "ep_07.4", "ep_class_07", "ep_07.1_explan", 
    "ep_08.1", "ep_class_08", "ep_08.1_explan", 
    "ep_08.2", "ep_class_08", "ep_08.1_explan", 
    "ep_08.3", "ep_class_08", "ep_08.1_explan", 
    "ep_08.4", "ep_class_08", "ep_08.1_explan", 
    "ep_08.5", "ep_class_08", "ep_08.5_explan", 
    "ep_09.1", "ep_class_09", "ep_09.1_explan", 
    "ep_09.2", "ep_class_09", "ep_09.1_explan", 
    "ep_09.3", "ep_class_09", "ep_09.1_explan", 
    "ep_101", "ep_class_10", "ep_101_explan", 
    "ep_102", "ep_class_10", "ep_101_explan", 
    "ep_103", "ep_class_10", "ep_103_explan"
) %>% 
    mutate(ep_code = factor(ep_code),
           ep_class = factor(ep_class)) %>% 
    write_vc(ep_path,
             root,
             sorting = "ep_code",
             optimize = FALSE,
             strict = FALSE)
```


