# Collecting 'schemes' and 'scheme_types' information and moving data into three n2khab data sources

The concerned data sources are:

- `schemes` and `scheme_types` (to be created, uses either codes or English-only text)
- `namelist` (accommodates names in multiple languages, for variables of interest that are represented by codes)

## Information from the monitoring programme for the natural environment (MNE)

### Collecting information for `schemes` and `scheme_types`

Reading dataframe coming from the [n2khab-mne-selections](https://gitlab.com/florisvdh/n2khab-mne-selections) repo (a git LFS repo on Gitlab):

```{r}
schemes_types_nl <- 
    read_vc("rawraw_data/10_compmeetnet_types_milieudrukken")
```

Reading a link between older Dutch codes of environmental pressures and the current ep_codes:

```{r message=FALSE}
ws_list <- sheet_names(ep_gs_id)
ep_codes <- 
    ep_gs_id %>%
    read_sheet(sheet = ws_list[str_detect(ws_list, "\\ 1.0\\ ")]) %>% 
    select(ep_code = Environmental_pressure_code,
           ep_code_nl = Environmental_pressure_code_NL_OLD) %>% 
    mutate(ep_code = factor(ep_code,
                            levels = 
                                read_vc(ep_path, root) %>% 
                                .$ep_code %>% 
                                levels),
           ep_code_nl = plyr::mapvalues(ep_code,
                                        from = ep_code,
                                        to = ep_code_nl)
           )
```

Making a dataframe that combines the information of `schemes` and `scheme_types`:

```{r}
types <- read_vc(types_path, root)
# for 5.1, types with HC23 must be doubled:
types_hc <-
  types %>% 
  select(type, hydr_class) %>% 
  mutate(type = as.character(type),
         second_class = ifelse(hydr_class == "HC23",
                               "HC2", 
                               NA),
         hydr_class = ifelse(hydr_class == "HC23",
                             "HC3", 
                             as.character(hydr_class))) %>% 
  pivot_longer(cols = contains("class"),
               names_to = "dummy",
               values_to = "hydr_class_expand",
               values_drop_na = TRUE) %>% 
  select(-dummy)
schemes_schemetypes <- 
    schemes_types_nl %>% 
    select(attribute_1 = Comp_meetnet,
           attribute_2 = Druk,
           tag_1 = MaxMilieudrukkengroep,
           type = Vegcode) %>% 
    mutate(type = as.character(type)) %>% 
    inner_join(types_hc, by = "type") %>% 
    mutate(
        programme = "MNE" %>% factor,
        attribute_1 = recode(attribute_1,
                             "Atmosferisch meetnet" = "ATM",
                             "Bodemmeetnet" = "SOIL",
                             "Grondwatermeetnet" = "GW",
                             "Oppervlaktewatermeetnet" = "SURF",
                             "Inundatiemeetnet" = "INUN"),
        attribute_2 = factor(attribute_2,
                             levels = 
                                 ep_codes %>% 
                                 .$ep_code_nl %>% 
                                 levels) %>% 
                        plyr::mapvalues(.,
                            from = levels(.),
                            to = ep_codes %>% 
                                 .$ep_code %>% 
                                 levels
                        ),
        attribute_3 = ifelse(attribute_1 == "GW" & attribute_2 == "ep_05.1",
                             ifelse(hydr_class_expand == "HC3",
                                    "aq",
                                    ifelse(type == "8310",
                                           "quarries",
                                           "terr")
                                    ),
                             ifelse(attribute_1 == "SURF" & 
                                      attribute_2 == "ep_03.4",
                                    ifelse(type %in% c("3260", "7220"),
                                           "lotic",
                                           "lentic"),
                                    NA)
                             ) %>% 
                      factor,
        scheme = str_c(attribute_1, 
                       str_match(attribute_2, "ep(_.+)")[,2],
                       ifelse(is.na(attribute_3),
                              "",
                              str_c("_",
                                    attribute_3))
                       ) %>% 
                 factor,
        tag_1 = ifelse(tag_1 == "a",
                       "focal",
                       "secondary") %>% 
                factor,
        type = factor(type,
                      levels = 
                        types %>% 
                        .$type %>% 
                        levels)
    ) %>% 
    select(-hydr_class_expand) %>% 
    distinct() %>% 
    arrange(programme, scheme)
```


### Make the `schemes` data source with MNE content

Set path and filename of `schemes` and `scheme_types`:

```{r}
schemes_path <- "inst/textdata/schemes"
scheme_types_path <- "inst/textdata/scheme_types"
```

Separately defining spatial restrictions for specific schemes:

```{r}
spat_restr <- 
    tibble(
      scheme = c("GW_05.1_terr"),
      spatial_restriction = 
        paste("Restrict to zones with shallow groundwater",
              "(in reach of vegetation).",
              "Exclude aquatic (i.e. rivulet) forms of type 7220.",
              "Don't exclude locations with types that are",
              "'(almost) everywhere groundwater dependent' (GD2).")
    ) %>% 
    bind_rows(tibble(
      scheme = c("GW_03.3",
                 "GW_04.2", 
                 "GW_05.2", 
                 "GW_07.1", 
                 "GW_08.3"),
      spatial_restriction = 
        paste("For terrestrial types: restrict to zones with shallow groundwater",
              "(in reach of vegetation).",
              "Don't exclude locations with types that are",
              "'(almost) everywhere groundwater dependent' (GD2).")
    )
    ) %>% 
    bind_rows(tibble(
        scheme = c("INUN_062",
                   "INUN_07.2"),
        spatial_restriction = "Restrict to areas susceptible to flooding."
            )
        ) %>% 
    bind_rows(tibble(
        scheme = c("INUN_03.4", 
                   "INUN_061",
                   "INUN_08.4"),
        spatial_restriction = 
          paste("Restrict to areas susceptible to flooding.",
                "Exclude aquatic (i.e. rivulet) forms of type 7220, if applicable."
                )
            )
        ) %>% 
    bind_rows(tibble(
        scheme = c("INUN_063",
                   "INUN_064",
                   "INUN_065",
                   "INUN_066"),
        spatial_restriction = 
          "Exclude aquatic (i.e. rivulet) forms of type 7220, if applicable."
            )
        ) %>% 
    bind_rows(tibble(
        scheme = c("SURF_012", 
                   "SURF_03.2", 
                   "SURF_03.4_lotic", 
                   "SURF_064", 
                   "SURF_066", 
                   "SURF_08.2", 
                   "SURF_08.4"),
        spatial_restriction = 
          "Exclude terrestrial (i.e. mire) forms of type 7220, if applicable."
            )
        ) %>% 
    bind_rows(tibble(
        scheme = "GW_05.1_aq",
        spatial_restriction = 
          paste("Exclude terrestrial (i.e. mire) forms of type 7220, if applicable.",
                "Exclude locations where the surface water is not connected with (nor affected by) the phreatic aquifer.")
            )
        ) %>% 
    bind_rows(tibble(
        scheme = "GW_05.1_quarries",
        spatial_restriction = 
          "Restrict to 8310 marl quarries that are directly connected with (intersected by) a groundwater aquifer."
            )
        ) %>% 
    mutate(scheme = factor(scheme,
                           levels = 
                               schemes_schemetypes %>% 
                               .$scheme %>% 
                               levels))
```

Preparing `schemes` with MNE content:

```{r}
schemes <- 
  schemes_schemetypes %>% 
    distinct(scheme,
             programme,
             attribute_1,
             attribute_2,
             attribute_3,
             tag_1) %>% 
    left_join(spat_restr, by = "scheme") %>% 
    mutate(notes = NA %>% as.character,
           tag_2 = NA %>% as.character,
           tag_3 = NA %>% as.character
           ) %>% 
    select(scheme,
           programme,
           attribute_1,
           attribute_2,
           attribute_3,
           spatial_restriction,
           notes,
           tag_1,
           tag_2,
           tag_3) %>% 
    arrange(programme, scheme)
```

### Make the `scheme_types` data source with MNE content

Collecting typegroup information to join on type:

```{r message=FALSE}
ws_list <- sheet_names(tg_gs_id)

typegroups <- 
    
    tg_gs_id %>%
    read_sheet(sheet = ws_list[str_detect(ws_list, "Eutrofiëring")],
               .name_repair = make.names) %>% 
    slice(2:n()) %>% 
    expand_grid(scheme = c("GW_03.3",
                           "SURF_03.4_lentic")
                ) %>% 
    select(scheme,
           type = Type,
           typegroup = Cluster.voedselrijkdom) %>% 
    
    bind_rows(
        tg_gs_id %>%
        read_sheet(sheet = "GW_05.1 (XG3)",
                   .name_repair = make.names) %>% 
        filter(Type %in% 
                 (schemes_schemetypes %>% 
                    filter(scheme == "GW_05.1_terr") %>% 
                    .$type)
               ) %>% 
        mutate(scheme = "GW_05.1_terr") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    bind_rows(
        tg_gs_id %>%
        read_sheet(sheet = "GW_05.1_aq (XG3)",
                   .name_repair = make.names) %>%
        filter(Type %in% 
                 (schemes_schemetypes %>% 
                    filter(scheme == "GW_05.1_aq") %>% 
                    .$type)
               ) %>% 
        mutate(scheme = "GW_05.1_aq") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    bind_rows(
        tg_gs_id %>%
        read_sheet(sheet = ws_list[str_detect(ws_list, "ATM_03\\.1")],
                   .name_repair = make.names) %>% 
        mutate(scheme = "ATM_03.1") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    bind_rows(
        tg_gs_id %>%
        read_sheet(sheet = ws_list[str_detect(ws_list, "ATM_04\\.1")],
                   .name_repair = make.names) %>% 
        mutate(scheme = "ATM_04.1") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    mutate(scheme = factor(scheme, 
                           levels = 
                             schemes_schemetypes$scheme %>% 
                             levels),
           type = factor(type, 
                         levels = 
                             schemes_schemetypes$type %>% 
                             levels),
           typegroup = str_c(scheme, 
                             "_group", 
                             typegroup) %>% 
                        factor
           )
```

Preparing `scheme_types` with MNE content:

```{r}
scheme_types <- 
  schemes_schemetypes %>% 
    distinct(scheme,
             type) %>% 
    left_join(typegroups, 
              by = c("scheme", "type")) %>% 
    arrange(scheme, type)
```


### Append associated names (multi-language) to the `namelist` datasource

Appending MNE names for `attribute_1`, `attribute_3`, `tag_1` and `programme`:

```{r}
cbind(code = c("ATM", "SOIL", "GW", "SURF", "INUN"),
       lang = c(rep("en", 5), rep("nl", 5))
       ) %>% 
    cbind(name = c("Atmospheric monitoring",
                   "Soil monitoring",
                   "Groundwater monitoring",
                   "Surfacewater monitoring",
                   "Inundationwater monitoring",
                   "Atmosferisch meetnet", 
                   "Bodemmeetnet", 
                   "Grondwatermeetnet", 
                   "Oppervlaktewatermeetnet", 
                   "Inundatiemeetnet")) %>% 
    rbind(matrix(c("aq", "en", "aquatic",
                   "terr", "en", "terrestrial",
                   "quarries", "en", "marl quarries",
                   "aq", "nl", "aquatisch",
                   "terr", "nl", "terrestrisch",
                   "quarries", "nl", "mergelgroeven",
                   "lotic", "en", "lotic waters",
                   "lentic", "en", "lentic waters",
                   "lotic", "nl", "stromende wateren",
                   "lentic", "nl", "stilstaande wateren"),
                 ncol = 3,
                 byrow = TRUE)) %>% 
    rbind(matrix(c("focal", "en", "Focal scheme",
                   "secondary", "en", "Secondary scheme",
                   "focal", "nl", "Hoofdmeetnet",
                   "secondary", "nl", "Secundair meetnet"),
                 nrow = 4,
                 byrow = TRUE)) %>% 
    rbind(matrix(c("MNE", "en", "Monitoring programme for the Natural Environment",
                   "MHQ", "en", "Monitoring programme for biotic Habitat Quality",
                   "MNE", "nl", "Monitoringprogramma natuurlijk milieu",
                   "MHQ", "nl", "Monitoringprogramma biotische habitatkwaliteit"),
                 nrow = 4,
                 byrow = TRUE)) %>% 
    as_tibble %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Appending MNE names and shortnames for `scheme`:

```{r message=FALSE}
namelist <- read_vc(namelist_path, root)
schemes %>% 
    filter(programme == "MNE") %>% 
    distinct(scheme, attribute_1, attribute_2, attribute_3) %>% 
    mutate_all(as.character) %>% 
    rename(code = scheme) %>% 
    mutate(code = as.character(code)) %>% 
    crossing(lang = c("en", "nl")) %>% 
    left_join(namelist %>% select(-shortname),
              by = c("attribute_1" = "code", "lang")) %>% 
    rename(name_1 = name) %>% 
    left_join(tibble(lang = c("en", "nl"),
                     name_2 = c("environmental pressure",
                                "milieudruk")),
              by = "lang") %>% 
    left_join(namelist %>% select(-name),
              by = c("attribute_2" = "code", "lang")) %>% 
    rename(name_3 = shortname) %>% 
    left_join(namelist %>% select(-shortname),
              by = c("attribute_3" = "code", "lang")) %>% 
    rename(name_4 = name) %>% 
    mutate(name = str_c(name_1, 
                        ": ",
                        name_2,
                        " ",
                        name_3,
                        ifelse(is.na(name_4),
                                 "",
                                 str_c(": partim ",
                                       name_4)
                                 )
                    ),
           shortname = str_c(attribute_1,
                             ": ",
                             name_3,
                             ifelse(is.na(attribute_3),
                                 "",
                                 str_c(": ",
                                       attribute_3)
                                 )
                             )
           ) %>% 
    select(code, lang, name, shortname) %>% 
    bind_rows(read_vc(namelist_path, root),
              .) %>% 
    write_vc(namelist_path, root)
```

Appending names and shortnames for typegroups of MNE schemes:

```{r}
tribble(~nr, ~lang, ~name, ~shortname,
        1, "en", "very sensitive non-forest types", "very sensitive non-forest",
        1, "nl", "zeer gevoelige niet-bostypes", "zeer gevoelig niet-bos",
        2, "en", "sensitive non-forest types", "sensitive non-forest",
        2, "nl", "gevoelige niet-bostypes", "gevoelig niet-bos",
        3, "en", "less to not sensitive non-forest types", "less/not sensitive non-forest",
        3, "nl", "minder tot niet gevoelige niet-bostypes", "minder/niet gevoelig niet-bos",
        4, "en", "very sensitive forest types", "very sensitive forest",
        4, "nl", "zeer gevoelige bostypes", "zeer gevoelig bos",
        5, "en", "sensitive forest types", "sensitive forest",
        5, "nl", "gevoelige bostypes", "gevoelig bos",
        6, "en", "less to not sensitive forest types", "less/not sensitive forest",
        6, "nl", "minder tot niet gevoelige bostypes", "minder/niet gevoelig bos"
        ) %>% 
  crossing(prefix = c("ATM_03.1_group", "ATM_04.1_group")) %>% 
  mutate(code = str_c(prefix, nr)) %>% 
  select(code, lang, name, shortname) %>% 
  bind_rows(
    tribble(~nr, ~lang, ~name, ~shortname,
        1, "en", "types from oligotrophic environments", "oligotrophic",
        1, "nl", "types van oligotroof milieu", "oligotroof",
        2, "en", "types from mesotrophic environments", "mesotrophic",
        2, "nl", "types van mesotroof milieu", "mesotroof",
        3, "en", "types from weakly eutrophic environments", "weakly eutrophic",
        3, "nl", "types van zwak eutroof milieu", "zwak eutroof",
        4, "en", "types from moderately eutrophic to eutrophic environments", "(moderately) eutrophic",
        4, "nl", "types van matig eutroof tot eutroof milieu", "(matig) eutroof"
        ) %>% 
      crossing(prefix = c("GW_03.3_group", "SURF_03.4_lentic_group")) %>% 
      mutate(code = str_c(prefix, nr)) %>% 
      select(code, lang, name, shortname)
  ) %>% 
  bind_rows(
        tribble(~code, ~lang, ~name, ~shortname,
        1, "en", "types from very wet environments", "very wet",
        1, "nl", "types van zeer nat milieu", "zeer nat",
        2, "en", "types from wet environments", "wet",
        2, "nl", "types van nat milieu", "nat",
        3, "en", "types from moderately wet environments", "moderately wet",
        3, "nl", "types van matig nat milieu", "matig nat",
        4, "en", "types from moist environments", "moist",
        4, "nl", "types van vochtig milieu", "vochtig",
        5, "en", "types from locally moist environments", "locally moist",
        5, "nl", "types van lokaal vochtig milieu", "lokaal vochtig"
        ) %>% 
      mutate(code = str_c("GW_05.1_terr_group", code))
  ) %>% 
  bind_rows(
        tribble(~code, ~lang, ~name, ~shortname,
        1, "en", "types from spring-fed brooks", "spring-fed brooks",
        1, "nl", "types van bronbeken", "bronbeken",
        2, "en", "types from lakes with stable groundwater table", "stable lakes",
        2, "nl", "types van stilstaand water met stabiel grondwaterregime", "stabiele plassen",
        3, "en", "types from lowland rivers", "lowland rivers",
        3, "nl", "types van laaglandbeken", "laaglandbeken",
        4, "en", "types from dune lakes (moderate groundwater dynamics)", "dune lakes",
        4, "nl", "types van duinplassen (middelmatig grondwaterregime)", "duinplassen",
        5, "en", "types from softwater lakes with dynamic groundwater table", "dynamic softwater lakes",
        5, "nl", "types van vennen met dynamisch grondwaterregime", "dynamische vennen"
        ) %>% 
      mutate(code = str_c("GW_05.1_aq_group", code))
  ) %>% 
  bind_rows(read_vc(namelist_path, root),
              .) %>% 
  write_vc(namelist_path, root)
```



## Appending records for the monitoring programme for biotic habitat quality (MHQ)

Writing `schemes`.

```{r message=FALSE}
schemes_MHQ <- 
  mhq_gs_id %>%
  read_sheet(sheet = 1, col_types = "c") %>%
  mutate(attribute_1 = ifelse(attribute_1 == 91, 
                              "91E0", 
                              as.character(attribute_1)),
         scheme = as.factor(scheme),
         programme = as.factor(programme),
         attribute_1 = as.factor(attribute_1),
         tag_1 = as.factor(tag_1)) 

schemes %>%
  bind_rows(schemes_MHQ) %>%
  mutate(scheme = factor(scheme,
                            levels = c(
                              levels(schemes$scheme),
                              levels(schemes_MHQ$scheme)
                              )
                            ),
         programme = factor(programme,
                            levels = c(
                              levels(schemes$programme),
                              levels(schemes_MHQ$programme)
                              )
                            ),
         attribute_1 = factor(attribute_1,
                            levels = c(
                              levels(schemes$attribute_1),
                              levels(schemes_MHQ$attribute_1)
                              )
                            ),
         tag_1 = factor(tag_1,
                           levels = c(
                              levels(schemes$tag_1),
                              levels(schemes_MHQ$tag_1)
                              )
                           )
         ) %>%
  arrange(programme, scheme) %>% 
    write_vc(schemes_path,
             root,
         sorting = c("programme", 
                     "scheme"),
         optimize = TRUE,
         strict = FALSE)

```

Writing `scheme_types`.

```{r}
scheme_type_MHQ <- 
  mhq_gs_id %>%
  read_sheet(sheet = 2) %>%
  mutate(scheme = factor(scheme),
         type = unlist(type))
  
scheme_types %>%
  bind_rows(scheme_type_MHQ) %>%
  mutate(scheme = factor(scheme,
                            levels = c(
                              levels(scheme_types$scheme),
                              levels(scheme_type_MHQ$scheme)
                              )),
         type = factor(type,
                       levels = levels(scheme_types$type))
  ) %>% 
  arrange(scheme, type) %>% 
  write_vc(scheme_types_path,
           root,
           sorting = c("scheme",
                     "type"),
           optimize = TRUE,
           strict = FALSE)
                         

```

Appending MHQ names for `scheme`.

```{r message=FALSE}

namelist <- read_vc(namelist_path, root)

schemes_MHQ_names <- schemes_MHQ %>%
  left_join(namelist, by = c("attribute_1" = "code")) %>%
  mutate(name = paste0(ifelse(lang == "en",
                             "Habitat quality scheme: ",
                             "Meetnet habitatkwaliteit: "),
                      shortname,
                      ifelse(scheme == "HQ2190_terr",
                             ifelse(lang == "en",
                                    " (terrestrial)", " (terrestrisch)"),
                             "")
                      ),
         shortname = paste0(ifelse(lang == "en",
                             "Habitat quality scheme ",
                             "Meetnet habitatkwaliteit "),
                      attribute_1,
                      ifelse(scheme == "HQ2190_terr",
                             " (terr)",
                             ""))
         ) %>%
  select(code = scheme, lang, name, shortname)

namelist %>%
  bind_rows(schemes_MHQ_names) %>%
  write_vc(namelist_path, root) 

```

