# Check textdata

## namelist

```{r}
namelist_check <- 
    read_vc(namelist_path, root)
```

Are all codes unique per language?

```{r}
namelist_check %>% 
    count(code, lang) %>% 
    filter(n > 1) %>% 
    nrow == 0
```

No lines without names or shortname?

```{r}
namelist_check %>% 
    filter(is.na(name), is.na(shortname)) %>% 
    nrow == 0
```

All lines have a value for code and lang?

```{r}
namelist_check %>% 
    filter(is.na(code) | is.na(lang)) %>% 
    nrow == 0
```


Language statistics:

```{r}
namelist_check %>% 
    mutate(name = !is.na(name),
           shortname = !is.na(shortname)) %>% 
    group_by(lang) %>% 
    summarize(name = sum(name),
              shortname = sum(shortname))
```


## types

```{r}
types_check <- read_vc(types_path, root)
```

Are all codes unique?

```{r}
types_check %>% 
    count(type) %>% 
    filter(n > 1) %>% 
    nrow == 0
```

Are all main types correct?

```{r}
types_check %>% 
    mutate(type = as.character(type),
           main_type_2 = ifelse(str_detect(type, "_|\\+"),
                                str_sub(type, 
                                        end = str_locate(type, "_|\\+") - 1),
                                type)
    ) %>% 
    filter(main_type != main_type_2) %>% 
    nrow == 0
```

Which are the frequencies of different combinations of hydrological class, groundwater and flood dependency?

```{r warning=FALSE}
types_check %>% 
    count(hydr_class, groundw_dep, flood_dep)
```

Inspecting the `types` data source, sorted according to the above three variables:

```{r}
types_check %>% 
    arrange(hydr_class, groundw_dep, flood_dep)
```

## env_pressures

```{r}
ep_check <- read_vc(ep_path, root)
```


Are all codes unique?

```{r}
ep_check %>% 
    count(ep_code) %>% 
    filter(n > 1) %>% 
    nrow == 0
```

Do ep_classes coincide between ep_code and ep_class?

```{r}
ep_check %>% 
    mutate(ep_class_c = str_c("ep_class_",
                        str_match(ep_code, "ep_(..).*")[,2]
                        ),
           ep_class = as.character(ep_class)
    ) %>% 
    (function(df) {
        all.equal(df$ep_class_c, df$ep_class)
    })
```

Do ep_classes coincide between ep_code and explanation?

```{r}
ep_check %>% 
    mutate(ep_class_c1 = str_match(ep_code, "ep_(..).*")[,2],
           ep_class_c2 = str_match(explanation, "ep_(..).*")[,2],
    ) %>% 
    (function(df) {
        all.equal(df$ep_class_c1, df$ep_class_c2)
    })
```

## schemes

```{r}
schemes_check <- 
    read_vc(schemes_path, root)
```


Are all scheme codes unique?

```{r}
schemes_check %>% 
    count(scheme) %>% 
    filter(n > 1) %>% 
    nrow == 0
```

Which are the represented programmes (projects)?

```{r}
schemes_check %>% 
    distinct(programme)
```

Are all `attribute_2` values of the MNE programme present in `env_pressures`?

```{r}
schemes_check %>% 
    filter(programme == "MNE") %>% 
    distinct(attribute_2) %>% 
    .$attribute_2 %in% ep_check$ep_code %>% 
    all
```


## scheme_types

```{r}
scheme_types_check <- 
    read_vc(scheme_types_path, root)
```


No double combinations occur?

```{r}
scheme_types_check %>% 
    count(scheme, type) %>% 
    filter(n > 1) %>% 
    nrow == 0
```

All occurring schemes are present in `schemes`?

```{r}
scheme_types_check %>% 
    .$scheme %in% schemes_check$scheme %>% 
    all
```

All occurring types are present in `types`?

```{r}
scheme_types_check %>% 
    .$type %in% types_check$type %>% 
    all
```

When a scheme-type combination has a typegroup, is this always the case for all types of the concerned scheme?

```{r message=FALSE}
scheme_types_check %>% 
    group_by(scheme) %>% 
    summarise(nr_entries = sum(!is.na(typegroup))) %>% 
    filter(nr_entries > 0) %>% 
    inner_join(scheme_types_check) %>% 
    filter(is.na(typegroup)) %>% 
    nrow == 0
```



