---
title: "Generating textdata files for the n2khab package"
subtitle: "Reference lists for scientific workflows regarding Natura 2000 habitat in Flanders"
date: "`r lubridate::now()`"
link-citations: true
linkcolor: link.colour
citecolor: link.colour
urlcolor: link.colour
geometry: margin=1in
mainfont: "Calibri"
fontsize: 11pt
documentclass: "article"
# csl: ../inbo.csl
# bibliography: ../references.bib
site: bookdown::bookdown_site
output:
  bookdown::html_document2:
    keep_md: TRUE
    number_sections: yes
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../header.html
  bookdown::pdf_document2:
    fig_caption: yes
    keep_tex: yes
    toc: yes
    toc_depth: 3
    latex_engine: xelatex
    number_sections: true
    includes:
        in_header: ../header.tex
---

```{r setup, include=FALSE}
renv::restore()
options(stringsAsFactors = FALSE)
# Setup for googlesheets4 authentication. Set the appropriate env vars in
# .Renviron and make sure you ran gs4_auth() interactively with these settings
# for the first run (or to renew an expired Oauth token)
if (Sys.getenv("GARGLE_OAUTH_EMAIL") != "") {
    options(gargle_oauth_email = Sys.getenv("GARGLE_OAUTH_EMAIL"))
}
if (Sys.getenv("GARGLE_OAUTH_CACHE") != "") {
    options(gargle_oauth_cache = Sys.getenv("GARGLE_OAUTH_CACHE"))
}
library(tidyverse)
library(stringr)
library(knitr)
library(googlesheets4)
local_gs4_quiet()
library(git2rdata)
opts_chunk$set(
  echo = TRUE,
  dpi = 300
)
```

**Note: this is a bookdown project, supposed to be run from within the `misc/generate_textdata` subfolder. You can use the `generate_textdata.Rproj` RStudio project file in this subfolder to run it, and run `bookdown::render_book("index.Rmd", "bookdown::html_document2")`. To run in a separate R session, use RStudio's build button, or (for older renv libraries) with R or Rscript (use the right version) in a shell.**

```{r}
# IDs of all involved googlesheets
types_gs_id <- "1dK0S1Tt3RlVEh4WrNF5N_-hn0OXiTUVOvsL2JRhSl78"
wdep_gs_id <- "1bhXgamK28K--MSWF7goKNOWtWPEI149_QCpU-3V_k_E"
ep_gs_id <- "1PH6InqJk0ijQF_N7v7IZjarijlqHRBCKhTbDn44skZU"
tg_gs_id <- "1n2ohvuLEK_anX37gxlanQXRGotweMVonFxKxtfTcuZI"
mhq_gs_id <- "1PU9MDyJKZz2DOrKqb-SIIVmSaR1mh4Ke6v0WCqhr8t4"
```




