# Collecting 'schemes' and 'scheme_types' information and moving data into three n2khabutils data sources

The concerned data sources are:

- `schemes` and `scheme_types` (to be created, uses either codes or English-only text)
- `namelist` (accommodates names in multiple languages, for variables of interest that are represented by codes)

## Information from the monitoring programme for the natural environment (MNE)

### Collecting information for `schemes` and `scheme_types`

Reading dataframe coming from the [n2khab-mne-selections](https://gitlab.com/florisvdh/n2khab-mne-selections) repo (a git LFS repo on Gitlab):

```{r}
schemes_types_nl <- 
    read_vc("rawraw_data/10_compmeetnet_types_milieudrukken")
```

Reading a link between older Dutch codes of environmental pressures and the current ep_codes:

```{r message=FALSE}
ws_list <- gs_ws_ls(ep_source)
ep_codes <- 
    ep_source %>%
    gs_read(ws = ws_list[str_detect(ws_list, "\\ 1.0\\ ")], 
            check.names = T,
            verbose = FALSE) %>% 
    select(ep_code = Environmental_pressure_code,
           ep_code_nl = Environmental_pressure_code_NL_OLD) %>% 
    mutate(ep_code = factor(ep_code,
                            levels = 
                                read_vc(ep_path) %>% 
                                .$ep_code %>% 
                                levels),
           ep_code_nl = plyr::mapvalues(ep_code,
                                        from = ep_code,
                                        to = ep_code_nl)
           )
```

Making a dataframe that combines the information of `schemes` and `scheme_types`:

```{r}
schemes_schemetypes <- 
    schemes_types_nl %>% 
    select(attribute_1 = Comp_meetnet,
           attribute_2 = Druk,
           attribute_3 = Waterhuishoudingsklasse_2,
           tag_1 = MaxMilieudrukkengroep,
           type = Vegcode) %>% 
    mutate(
        programme = "MNE" %>% factor,
        attribute_1 = recode(attribute_1,
                             "Atmosferisch meetnet" = "ATM",
                             "Bodemmeetnet" = "SOIL",
                             "Grondwatermeetnet" = "GW",
                             "Oppervlaktewatermeetnet" = "SURF",
                             "Inundatiemeetnet" = "INUN"),
        attribute_2 = factor(attribute_2,
                             levels = 
                                 ep_codes %>% 
                                 .$ep_code_nl %>% 
                                 levels) %>% 
                        plyr::mapvalues(.,
                            from = levels(.),
                            to = ep_codes %>% 
                                 .$ep_code %>% 
                                 levels
                        ),
        attribute_3 = ifelse(attribute_1 == "GW" & attribute_2 == "ep_05.1",
                             ifelse(attribute_3 == "Oppervlaktewater",
                                    "aq",
                                    "terr"),
                             NA) %>% 
                      factor,
        scheme = str_c(attribute_1, 
                       str_match(attribute_2, "ep(_.+)")[,2],
                       ifelse(is.na(attribute_3),
                              "",
                              str_c("_",
                                    attribute_3))
                       ) %>% 
                 factor,
        tag_1 = ifelse(tag_1 == "a",
                       "focal",
                       "secondary") %>% 
                factor,
        type = as.character(type) %>% 
                factor(levels = 
                           read_vc(types_path) %>% 
                           .$type %>% 
                           levels)
    ) %>% 
    arrange(programme, scheme)
```


### Make the `schemes` data source with MNE content

Set path and filename of `schemes` and `scheme_types`:

```{r}
schemes_path <- "../../n2khabutils/inst/textdata/schemes"
scheme_types_path <- "../../n2khabutils/inst/textdata/scheme_types"
```

Separately defining spatial restrictions for specific schemes:

```{r}
spat_restr <- 
    tribble(
        ~scheme, ~spatial_restriction,
        "GW_05.1_terr", "zones with shallow groundwater (in reach of vegetation)"
    ) %>% 
    bind_rows(tibble(
        scheme = c("INUN_03.4", 
                   "INUN_061",
                   "INUN_062",
                   "INUN_07.2",
                   "INUN_08.4"),
        spatial_restriction = "areas susceptible to flooding"
            )
        ) %>% 
    mutate(scheme = factor(scheme,
                           levels = 
                               schemes_schemetypes %>% 
                               .$scheme %>% 
                               levels))
```

Writing `schemes`:

```{r message=FALSE}
schemes_schemetypes %>% 
    distinct(scheme,
             programme,
             attribute_1,
             attribute_2,
             attribute_3,
             tag_1) %>% 
    left_join(spat_restr) %>% 
    mutate(notes = NA %>% as.character,
           tag_2 = NA %>% as.character,
           tag_3 = NA %>% as.character
           ) %>% 
    select(scheme,
           programme,
           attribute_1,
           attribute_2,
           attribute_3,
           spatial_restriction,
           notes,
           tag_1,
           tag_2,
           tag_3) %>% 
    arrange(programme, scheme) %>% 
    write_vc(schemes_path,
         sorting = c("programme", 
                     "scheme"),
         optimize = TRUE,
         strict = FALSE)
```

### Make the `scheme_types` data source with MNE content

Collecting typegroup information to join on type:

```{r message=FALSE}
tg_source <- gs_key("1n2ohvuLEK_anX37gxlanQXRGotweMVonFxKxtfTcuZI")
ws_list <- gs_ws_ls(tg_source)

typegroups <- 
    
    tg_source %>%
    gs_read(ws = ws_list[str_detect(ws_list, "Eutrofiëring")], 
            check.names = T,
            verbose = FALSE) %>% 
    slice(2:n()) %>% 
    mutate(scheme = "GW_03.3") %>% 
    select(scheme,
           type = Type,
           typegroup = Cluster.voedselrijkdom) %>% 
    
    bind_rows(
        tg_source %>%
        gs_read(ws = ws_list[str_detect(ws_list, "GW_05\\.1")], 
                check.names = T,
                verbose = FALSE) %>% 
        filter(!Type %in% 
                   (schemes_schemetypes %>% 
                        filter(scheme == "GW_05.1_aq") %>% 
                        .$type)
               ) %>% 
        mutate(scheme = "GW_05.1_terr") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    bind_rows(
        tg_source %>%
        gs_read(ws = ws_list[str_detect(ws_list, "ATM_03\\.1")], 
                check.names = T,
                verbose = FALSE) %>% 
        mutate(scheme = "ATM_03.1") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    bind_rows(
        tg_source %>%
        gs_read(ws = ws_list[str_detect(ws_list, "ATM_04\\.1")], 
                check.names = T,
                verbose = FALSE) %>% 
        mutate(scheme = "ATM_04.1") %>% 
        select(scheme,
               type = Type,
               typegroup = Typegroep)
    ) %>% 
    
    mutate(scheme = factor(scheme, 
                           levels = 
                             schemes_schemetypes$scheme %>% 
                             levels),
           type = factor(type, 
                         levels = 
                             schemes_schemetypes$type %>% 
                             levels),
           typegroup = str_c(scheme, 
                             "_group", 
                             typegroup) %>% 
                        factor
           )
```

Writing `scheme_types`:

```{r message=FALSE}
schemes_schemetypes %>% 
    distinct(scheme,
             type) %>% 
    left_join(typegroups) %>% 
    arrange(scheme, type) %>% 
    write_vc(scheme_types_path,
         sorting = c("scheme",
                     "type"),
         optimize = TRUE,
         strict = FALSE)
```


### Append associated names (multi-language) to the `namelist` datasource

Appending MNE names for `attribute_1`, `attribute_3`, `tag_1` and `programme`:

```{r}
cbind(code = c("ATM", "SOIL", "GW", "SURF", "INUN"),
       lang = c(rep("en", 5), rep("nl", 5))
       ) %>% 
    cbind(name = c("Atmospheric monitoring",
                   "Soil monitoring",
                   "Groundwater monitoring",
                   "Surfacewater monitoring",
                   "Inundationwater monitoring",
                   "Atmosferisch meetnet", 
                   "Bodemmeetnet", 
                   "Grondwatermeetnet", 
                   "Oppervlaktewatermeetnet", 
                   "Inundatiemeetnet")) %>% 
    rbind(matrix(c("aq", "en", "aquatic",
                   "terr", "en", "terrestrial",
                   "aq", "nl", "aquatisch",
                   "terr", "nl", "terrestrisch"),
                 nrow = 4,
                 byrow = TRUE)) %>% 
    rbind(matrix(c("focal", "en", "Focal scheme",
                   "secondary", "en", "Secondary scheme",
                   "focal", "nl", "Hoofdmeetnet",
                   "secondary", "nl", "Secundair meetnet"),
                 nrow = 4,
                 byrow = TRUE)) %>% 
    rbind(matrix(c("MNE", "en", "Monitoring programme for the Natural Environment",
                   "MHQ", "en", "Monitoring programme for biotic Habitat Quality",
                   "MNE", "nl", "Monitoringprogramma natuurlijk milieu",
                   "MHQ", "nl", "Monitoringprogramma biotische habitatkwaliteit"),
                 nrow = 4,
                 byrow = TRUE)) %>% 
    as_tibble %>% 
    mutate(shortname = NA %>% as.character) %>% 
    bind_rows(read_vc(namelist_path),
              .) %>% 
    write_vc(namelist_path)
```

Appending MNE names and shortnames for `scheme`:

```{r message=FALSE}
namelist <- read_vc(namelist_path)
read_vc(schemes_path) %>% 
    filter(programme == "MNE") %>% 
    distinct(scheme, attribute_1, attribute_2, attribute_3) %>% 
    mutate_all(as.character) %>% 
    rename(code = scheme) %>% 
    mutate(code = as.character(code)) %>% 
    crossing(lang = c("en", "nl")) %>% 
    left_join(namelist %>% select(-shortname),
              by = c("attribute_1" = "code", "lang")) %>% 
    rename(name_1 = name) %>% 
    left_join(tibble(lang = c("en", "nl"),
                     name_2 = c("environmental pressure",
                                "milieudruk"))) %>% 
    left_join(namelist %>% select(-name),
              by = c("attribute_2" = "code", "lang")) %>% 
    rename(name_3 = shortname) %>% 
    left_join(namelist %>% select(-shortname),
              by = c("attribute_3" = "code", "lang")) %>% 
    rename(name_4 = name) %>% 
    mutate(name = str_c(name_1, 
                        ": ",
                        name_2,
                        " ",
                        name_3,
                        ifelse(is.na(name_4),
                                 "",
                                 str_c(": partim ",
                                       name_4)
                                 )
                    ),
           shortname = str_c(attribute_1,
                             ": ",
                             name_3,
                             ifelse(is.na(attribute_3),
                                 "",
                                 str_c(":  ",
                                       attribute_3)
                                 )
                             )
           ) %>% 
    select(code, lang, name, shortname) %>% 
    bind_rows(read_vc(namelist_path),
              .) %>% 
    write_vc(namelist_path)
```



## Appending records for the monitoring programme for biotic habitat quality (MHQ)

Writing `schemes`.

```{r message=FALSE}

schemes_MHQ <- gs_key("1Lltpq1_OQENZKrCmfDLuBiaZQ0_YCoGkFAqcOtFRdHg") %>%
  gs_read(ws = 1) %>%
  mutate(attribute_1 = ifelse(attribute_1 == 91, 
                              "91E0", 
                              as.character(attribute_1)),
         scheme = as.factor(scheme),
         programme = as.factor(programme),
         attribute_1 = as.factor(attribute_1),
         tag_1 = as.factor(tag_1)) 

schemes <- read_vc(schemes_path)

schemes %>%
  bind_rows(schemes_MHQ) %>%
  mutate(scheme = factor(scheme,
                            levels = c(
                              levels(schemes$scheme),
                              levels(schemes_MHQ$scheme)
                              )
                            ),
         programme = factor(programme,
                            levels = c(
                              levels(schemes$programme),
                              levels(schemes_MHQ$programme)
                              )
                            ),
         attribute_1 = factor(attribute_1,
                            levels = c(
                              levels(schemes$attribute_1),
                              levels(schemes_MHQ$attribute_1)
                              )
                            ),
         tag_1 = factor(tag_1,
                           levels = c(
                              levels(schemes$tag_1),
                              levels(schemes_MHQ$tag_1)
                              )
                           )
         ) %>%
  arrange(programme, scheme) %>% 
    write_vc(schemes_path,
         sorting = c("programme", 
                     "scheme"),
         optimize = TRUE,
         strict = FALSE)

```

Writing `schemesTypes`.

```{r}
schemes_type <- read_vc(scheme_types_path)

scheme_type_MHQ <- gs_key("1Lltpq1_OQENZKrCmfDLuBiaZQ0_YCoGkFAqcOtFRdHg") %>%
  gs_read(ws = 2) %>%
  mutate(scheme = factor(scheme))
  
schemes_type %>%
  bind_rows(scheme_type_MHQ) %>%
  mutate(scheme = factor(scheme,
                            levels = c(
                              levels(schemes_type$scheme),
                              levels(scheme_type_MHQ$scheme)
                              )),
         type = factor(type,
                       levels = levels(schemes_type$type))
  ) %>% 
  arrange(scheme, type) %>% 
  write_vc(scheme_types_path,
           sorting = c("scheme",
                     "type"),
           optimize = TRUE,
           strict = FALSE)
                         

```

Appending MHQ names for `scheme`.

```{r message=FALSE}

namelist <- read_vc(namelist_path)

schemes_MHQ_names <- schemes_MHQ %>%
  left_join(namelist, by = c("attribute_1" = "code")) %>%
  mutate(name = paste(ifelse(lang == "en",
                             "Habitat quality scheme:",
                             "Meetnet habitatkwaliteit:"),
                      shortname),
         shortname = paste(ifelse(lang == "en",
                             "Habitat quality scheme",
                             "Meetnet habitatkwaliteit"),
                      attribute_1)) %>%
  select(code = scheme, lang, name, shortname)

namelist %>%
  bind_rows(schemes_MHQ_names) %>%
  write_vc(namelist_path) 

```

